---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# steg

<!-- badges: start -->
<!-- badges: end -->

The R package `steg` provides cryptographic tools for hiding messages inside photos.

It was developed as a fun project and **should not be used where serious cryptography is required**.

## Installation

You can install the development version of steg from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("DrJCampbell/steg")
```

## How this works

There are two main functions available - `hide_message()` and `reveal_message()`. These functions are convenience wrappers around a set of functions that together, process image data to store a message in the least-significant bit of a Portable Network Graphic (PNG) image file or reverse that process.

The PNG format is lossless, so data stored in a PNG file can be restored bit-for-bit as it were, prior to encoding.

The least significant bit of a pixel in a PNG image file has an imperceptible effect on how human eyes view the image. Because of this, we can modify the least-significant bit of an image without it having a noticible effect to the human eye. We can consider this image to be the host of the message we want to hide

We can create an image file where a message in text format has been written in monochrome. The bits of this image are either fully on, or fully off. We can take the binary on/off value of these bits and hide them in least significant bit of the host image without a human being able to see the change to the host image.

If we were to examine individual bits of an image where the least-significant bit encoded a message in image form, we would be able to see the message clearly. To help mitigate this possibility, we can encrypt the message bits with a key that is equal in length to the message images bits, using a Vernam cipher.

To encypt a message image with a key using a Vernam cipher, we use the `xor()` function on the corresponding image and message bits. This process is reversible, such that if we use the `xor()` function on the encrypted resulting bit with the corresponding kit bit, we will recover the original message bit. As long at the key bits are apparently random (and never reused), the Vernam cipher is _probably_ uncrackable.

To generate a stream of (hopefully enough) psuedo-random bits, we take a user-supplied key as text and use the SHA256 algorithm to create a hash of this key. We then use the resulting hash to create a new hash using the first hash result as the input. This hash result is appended to the existing hash. This process is repeated, with the result of the previous concatenated hashes used as the input for the next hash and appending to the previous results, until enough bits have been accumulated match the host image.

The least significant bit of the host image is replaced with the result of the `xor()` function using the binary value of the message-image bit and the bit of the corresponding bit-stream derived from the SHA256 cipher-block-chain (CBC) of the original key.


## Caveats

The effect of using a Vernam cipher on a stream of bits from a message and cryptographic key derived from a CBC of a key is fairly obvious compared to the original image. To explore this, a function called `show_image_layer()` is provided that allows visualisation of the bits at each position in an 8-bit image.


## Demonstration images

Two demonstration images are supplied with the package. `Hazels.PNG` is a four-channel image, meaning it has pixel arrays for the red, green, blue channels plus the alpha channel for transparency. `John_Tukey.png` is a one-chanel gray-scale image which is smalll in size (pixel dimensions). Thesen files can be accessed with:

`system.file("extdata", "Hazels.PNG", package = "steg")`

or 

`system.file("extdata", "John_Tukey.png", package = "steg")`


## Example

The simplest use of this package involves the following

```{r example_basic, eval=FALSE}
library(steg)

# Hide a message
hide_message(
  image=system.file("extdata", "Hazels.PNG", package = "steg"),
  #image=system.file("extdata", "John_Tukey.png", package = "steg"),
  key="Use_a_really_good_key_here",
  message="You can replace this message with up to 122 characters that will be written in a png file",
  crypto_image="crypto_image.png"
  )


# Reveal a message
reveal_message(
  key="Use_a_really_good_key_here",
  crypto_image="crypto_image.png")
```



You can also perform the individual steps used to hide or reveal a message using the following functions.


```{r example, eval=FALSE}

library(steg)

key = "Use_a_really_good_key_here"

# Using the Hazels.PNG file in the packages extdata folder
example_image_file = system.file("extdata", "Hazels.PNG", package = "steg")

# Get the bit-depth multiplier to apply to pixel values
bitdepth = get_image_bitdepth(file=example_image_file)

# Read the example image photo into a matrix of bits
image = read_original_image(file = example_image_file, bitdepth = bitdepth)

# Generate a pseudorandom string of bits using the sha256 algorithm seeded with the key
# The hashes are cipher-block-chained to produce as many bits as there are pixels in the image
key_bits = create_key_bits(key = key, image = image)

# Create an image containing the message to be hidden
message = create_message_image(
  file = "message.png",
  message = "You can replace this message with up to 122 characters that will be written in a png file",
  image = image
  )

# Hide the message in the image and write out as a png file
crypto = write_encrypted_image(
  "crypto.png", message, image, key = key
  )


# The crypto.png file could be sent to a third party. If they also know the key, they
# will be able to decrypt the message
plain_text = read_encrypted_image("crypto.png", plain_file = "decrypted.png", key = key)

# Plot the decrypted data to view the message
plot_decrypted_image(plain_text)
```


Finally, you can explore the individual bit-layers of any image using the following function.


```{r eval=FALSE}
show_image_layer(
  file='crypto.png',
  layer=1,
  channel=1
  )
```


```{r eval=FALSE}
show_image_layer(
  file=example_image_file,
  layer=1,
  channel=1
  )
```
